Research Proposal

### 2021-10-04 Reports/JOURNAL - Confidence Regions/ResearchProposal.tex
:-) :-) :+1:

\citet{ANewVersatileDiscreteDistribution} proposes a flexible discrete model, the ``discretised beta'' law.
The probability function of such a random variable is
\begin{equation}
\Pr(X  = x \mid \alpha, \beta, n_{\text{top}}, \zeta) = h(x) \exp\big\{ \alpha T_1(x) + \beta T_2(x) - A(\alpha, \beta) \big\},
\label{Eq:Probdb}
\end{equation}
where $\zeta=\text{TRUE}$ denotes that the support starts in zero ($n_{\text{bot}}=0$), otherwise it starts in one ($n_{\text{bot}}=1$), and
\begin{align}
x &\in \{n_{\text{bot}}, n_{\text{bot}}+1, \dots, n_{\text{top}}\},\\
h(x) & =\frac{(n_{\text{top}} - n_{\text{bot}}+2)^2}{
(x-n_{\text{bot}}+1)	(n_{\text{top}}-x+1)	},\\
T_1(x) &= \log\frac{x-n_{\text{bot}}+1}{n_{\text{top}}-n_{\text{bot}}+2}, \\
T_2(x) &= \log\frac{n_{\text{top}}-x+1}{n_{\text{top}}-n_{\text{bot}}+2}, \\
\intertext{and}
A(\alpha,\beta) &= \log \bigg(
\sum_{i=n_{\text{bot}}}^{n_{\text{top}}} h(i) \exp\big\{ \alpha T_1(i) + \beta T_2(i)\big\}
\bigg).
\end{align}
This definition yields a model that belongs to the exponential family for any $\alpha,\beta\in\mathbbm R$.
Denote this situation $X\sim\text{db}(\alpha,\beta, n_{\text{top}},\zeta)$.

If $\alpha,\beta>0$, we can relate the $\text{db}(\alpha,\beta, n_{\text{top}},\zeta)$ model with the (continuous) Beta distribution in the following manner:
\begin{equation}
\Pr(X=x \mid \alpha, \beta, n_{\text{top}}, \zeta) = 
\frac{1}{\kappa} f\Big(
\frac{x-n_{\text{bot}}+1}{n_{\text{top}}-n_{\text{bot}}+2}
\Big),
\label{Eq:Probdb}
\end{equation}
where
$$
\kappa = \sum_{i=n_{\text{bot}}}^{n_{\text{bot}}} f\Big(
\frac{x-n_{\text{bot}}+1}{n_{\text{top}}-n_{\text{bot}}+2}
\Big),
$$
and $f$ is the density of a Beta distribution with parameters $\alpha$ and $\beta$.

@@

We do not have that information, but we may approach the problem as an inferential one.
Let $S$ be any subset of the $H\times C$ plane.
We need 
\begin{equation}
\Pr\big((\mathfrak{H},\mathfrak{C})\in S\big)
\label{Eq:Problem}
\end{equation}
for all possible combinations of the sequence length $T$, embedding dimension $D$, and time delay $\tau$.
Since we do not know the distribution of $(\mathfrak{H},\mathfrak{C})$, we will model it.

Denote $\mathcal H$ the operator that transforms $\mathcal X$ into its histogram $\bm h$.
We want to know the joint distribution of $(\mathfrak{H}(\mathcal H(X)),\mathfrak{C}(\mathcal H(X)))$, where $X$ models the sequence $\mathcal X$.
Unfortunately, even knowing the distribution of $X$ under the null hypothesis (a collection of iid random variables), modelling $\mathcal H(X)$ is difficult.

All the patterns $(\pi_i)_{1\leq i\leq D!}$ are equally probable in the ideal white noise sequence.
We then use the $\text{db}(\alpha,\beta)$ model, and set $\alpha=\beta=1$, $n_{\text{top}}=D!$, and $\zeta=\text{FALSE}$ to describe $\mathcal H(X)$ under white noise.
With this, \eqref{Eq:Probdb} becomes $\Pr(\mathcal H(X)=\bm h)=(1/D!)_{1\leq i\leq D!}$ and, most importantly, given a histogram of patterns $\bm h$ we can estimate the model which produced it.
Denote this situation $\mathcal H(X)\sim\text{db}_0$.

Questions: given that $\mathcal H(X)\sim\text{db}_0$, what is the joint distribution of $(\mathfrak{H}(\mathcal H(X)),\mathfrak{C}(\mathcal H(X)))$?

Assume that for each combination of $(T,D,\tau)$, we have access to $N_{T,D,\tau}$ (denoted as $N$ in the following) independent identically distributed sequences $\bm x_1,\bm x_2, \dots, \bm x_N$ of white noise.
Each of these sequences produces a point in the $H\times C$ plane, namely $(h_1,c_1), (h_2,c_2), \dots , (h_N,c_N)$.
But let's remind how these points are produced.

The Bandt-Pompe symbolization transforms each sequence of values $\bm x_i$ into a sequence of symbols $\bm \pi_i$.
We build a histogram of proportions $\bm h_i$ with the sequence of symbols $\bm \pi_i$.
Now we assume that $\bm h_i$ is an outcome of size $N$ of $X\sim\text{db}(\alpha,\beta)$, and we compute $\widehat\alpha_i,\widehat\beta_i$.
In this way, the sequences $\bm x_1,\bm x_2, \dots, \bm x_N$ become histograms $\bm h_1, \bm h_2, \dots, \bm h_N$ and, on the one hand, they become points $(h_1,c_1), (h_2,c_2), \dots , (h_N,c_N)$ and estimates $(\widehat\alpha_1,\widehat\beta_1), (\widehat\alpha_2,\widehat\beta_2), \dots , (\widehat\alpha_N,\widehat\beta_N)$.
We may then use $\text{db}(\overline{\widehat{\alpha}_i},\overline{\widehat{\beta}_i})$ distribution as a model for the histograms.
From there, we may obtain an estimate of the distribution of $(\mathfrak{H},\mathfrak{C})$.

The problem was knowing $\Pr\big((\mathfrak{H},\mathfrak{C})\in S\big)$, but $\mathfrak{H}=\mathfrak{H}(X)$  and $\mathfrak{C} = \mathfrak{C}(X)$, and we may assume that $X\sim\text{db}(\alpha,\beta)$, and we may estimate $(\alpha,\beta)$.
The problem becomes finding the distribution of $\big(\mathfrak{H}(X),\mathfrak{C}(X)\big)$ under $X\sim\text{db}(\alpha,\beta)$.

>>>>

\citet{ANewVersatileDiscreteDistribution} proposes a flexible discrete model, the ``discretised beta'' law.
The probability function of such a random variable is
\begin{equation}
\Pr(X  = x \mid \alpha, \beta, n_{\text{top}}, \zeta) = h(x) \exp\big\{ \alpha T_1(x) + \beta T_2(x) - A(\alpha, \beta) \big\},
\label{Eq:ProbdbExponential}
\end{equation}
where $\zeta=\text{TRUE}$ denotes that the support starts in zero ($n_{\text{bot}}=0$), otherwise it starts in one ($n_{\text{bot}}=1$), and
\begin{align}
x &\in \{n_{\text{bot}}, n_{\text{bot}}+1, \dots, n_{\text{top}}\},\\
h(x) & =\frac{(n_{\text{top}} - n_{\text{bot}}+2)^2}{
(x-n_{\text{bot}}+1)	(n_{\text{top}}-x+1)	},\\
T_1(x) &= \log\frac{x-n_{\text{bot}}+1}{n_{\text{top}}-n_{\text{bot}}+2}, \\
T_2(x) &= \log\frac{n_{\text{top}}-x+1}{n_{\text{top}}-n_{\text{bot}}+2}, \\
\intertext{and}
A(\alpha,\beta) &= \log \bigg(
\sum_{i=n_{\text{bot}}}^{n_{\text{top}}} h(i) \exp\big\{ \alpha T_1(i) + \beta T_2(i)\big\}
\bigg).
\end{align}
This definition yields a model that belongs to the exponential family for any $\alpha,\beta\in\mathbbm R$.
Denote this situation $X\sim\text{db}(\alpha,\beta, n_{\text{top}},\zeta)$.

If $\alpha,\beta>0$, we can relate the $\text{db}(\alpha,\beta, n_{\text{top}},\zeta)$ model with the (continuous) Beta distribution in the following manner:
\begin{equation}
\Pr(X=x \mid \alpha, \beta, n_{\text{top}}, \zeta) = 
\frac{1}{\kappa} f\Big(
\frac{x-n_{\text{bot}}+1}{n_{\text{top}}-n_{\text{bot}}+2}
\Big),
\label{Eq:Probdb}
\end{equation}
where
$$
\kappa = \sum_{i=n_{\text{bot}}}^{n_{\text{bot}}} f\Big(
\frac{x-n_{\text{bot}}+1}{n_{\text{top}}-n_{\text{bot}}+2}
\Big),
$$
and $f$ is the density of a Beta distribution with parameters $\alpha$ and $\beta$, i.e.,
\begin{equation}
	f(u;\alpha, \beta) = \frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\Gamma(\beta)} u^{\alpha-1} (1-u)^{\beta-1} \mathbbm 1_{(0,1)}(u).
	\label{Eq:DensBeta}
\end{equation}
Denote this situation $U\sim\text{Beta}(\alpha,\beta)$.

@@

\begin{itemize}
\item[Q1:] What is the distribution of $(\mathfrak{H},\mathfrak{C})$ under $H_0:\mathcal X\text { is white noise}$?
\begin{itemize}
	\item $\bm X = (X_1,X_2,\dots,X_N)$ is white noise
	\item Fix $\tau=1$ and define $D$
	\item $\bm \pi=(\pi_1,\pi_2,\dots,\pi_{n-D})$ is the random sequence of patterns from $\bm X$. There are $D!$ possible patterns.
	\item The patterns are not independent because some cannot follow others, but we may shuffle them to avoid such restriction.
	\item $\bm h$ is the random histogram of proportions from $\bm \pi$
	\item $(\mathfrak{H},\mathfrak{C})$ is the random point in the Entropy-Complexity plane computed from $\bm h$
	\begin{itemize}	
		\item $H^*$ is the operator that transforms the random vector $\bm h$ into the random variable $\mathfrak{H}$
		\item $C^*$ is the operator that transforms the random vector $\bm h$ in to the random variable $\mathfrak{C}$
		\item Notice that $H^*$ does not compute the entropy of $\bm h$, which is a fixed value; it produces a random variable from the random variable $\bm h$. The same holds for $C^*$; it does not compute the complexity.
	\end{itemize}
\end{itemize}
\item[Q1:] What is the distribution of $
	(\mathfrak{H},\mathfrak{C}) = 
	\big(H^*(\bm h),C^*(\bm h)\big)$? Under $H_0$, $\bm h$ is (?) a collection of iid random variables taking values in $\{1, 2,\dots, D!\}$ with same probability $1/D!$ but, as they form a histogram, there must be a restriction on the sum of their values. I don't know how to go on from this point. Multinomial? Whichever the distribution of $\bm h$, how do we compute the distribution of $H^*(\bm h)$?
\item[$\bullet$] Alternatively, instead of considering such a flexible discrete distribution, we may use a parametric model.
\item[Q2:] What is the distribution of $
(\mathfrak{H},\mathfrak{C}) = 
\big(H^*(\bm h),C^*(\bm h)\big)$ when $\bm h\sim\text{db}(\alpha,\beta,1, D!)$?
\item[] But we go back to a quantity, rather than to a random variable.
\item[Q3:] Assume we have been able to compute the quantities $H^*(\alpha,\beta)$ and $C^*(\alpha,\beta)$ under the $\text{db}(\alpha,\beta,1, D!)$ model.
\item[] We have the maximum likelihood estimator for $(\alpha,\beta)$, say $(\widehat\alpha,\widehat\beta)$.
\item[] We know that the asymptotic distribution of $(\widehat\alpha,\widehat\beta)$ is a bivariate Gaussian law.
\item[] How close are we to finding an approximate distribution for $
(\mathfrak{H},\mathfrak{C})$ under $H_0$?
\end{itemize}

If $U\sim\text{Beta}(\alpha,\beta)$, then its entropy is
\begin{equation}
H(U) = \log\text{B}(\alpha,\beta) -
(\alpha-1)\psi(\alpha) -(\beta-1) \psi(\beta) +(\alpha+\beta-2) \psi(\alpha+\beta),
\label{Eq:HBeta}
\end{equation}
in which we denoted
$$
\text{B}(\alpha,\beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha+\beta)}.
$$

Also, if $U_1\sim\text{Beta}(\alpha_1,\beta_1)$ and
$U_2\sim\text{Beta}(\alpha_2,\beta_2)$, then the Kullback-Leibler divergence between the models is
\begin{multline}
D_{\text{KL}}(U_1,U_2) = \log\frac{\text{B}(\alpha_2,\beta_2)}{\text{B}(\alpha_1,\beta_1)} +
(\alpha_2-\alpha_1) \psi(\alpha) +
(\beta_2-\beta_1) \psi(\beta) + \\
(\alpha_2-\alpha_1+\beta_2-\beta_1) \psi(\alpha_1+\beta_1).
\label{Eq:DKLBeta}
\end{multline}
With~\eqref{Eq:DKLBeta} we may obtain the Jensen-Shannon distance between $U\sim\text{Beta}(\alpha,\beta)$ and the Uniform distribution that is $V\sim\text{Beta}(1,1)$.
The product between~\eqref{Eq:HBeta} and this Jensen-Shannon distance is the complexity.

%We do not have that information, but we may approach the problem as an inferential one.
%Let $S$ be any subset of the $H\times C$ plane.
%We need 
%\begin{equation}
%\Pr\big((\mathfrak{H},\mathfrak{C})\in S\big)
%\label{Eq:Problem}
%\end{equation}
%for all possible combinations of the sequence length $T$, embedding dimension $D$, and time delay $\tau$.
%Since we do not know the distribution of $(\mathfrak{H},\mathfrak{C})$, we will model it.
%
%Denote $\mathcal H$ the operator that transforms $\mathcal X$ into its histogram $\bm h$.
%We want to know the joint distribution of $(\mathfrak{H}(\mathcal H(X)),\mathfrak{C}(\mathcal H(X)))$, where $X$ models the sequence $\mathcal X$.
%Unfortunately, even knowing the distribution of $X$ under the null hypothesis (a collection of iid random variables), modelling $\mathcal H(X)$ is difficult.
%
%All the patterns $(\pi_i)_{1\leq i\leq D!}$ are equally probable in the ideal white noise sequence.
%We then use the $\text{db}(\alpha,\beta)$ model, and set $\alpha=\beta=1$, $n_{\text{top}}=D!$, and $\zeta=\text{FALSE}$ to describe $\mathcal H(X)$ under white noise.
%With this, \eqref{Eq:Probdb} becomes $\Pr(\mathcal H(X)=\bm h)=(1/D!)_{1\leq i\leq D!}$ and, most importantly, given a histogram of patterns $\bm h$ we can estimate the model which produced it.
%Denote this situation $\mathcal H(X)\sim\text{db}_0$.
%
%Questions: given that $\mathcal H(X)\sim\text{db}_0$, what is the joint distribution of $(\mathfrak{H}(\mathcal H(X)),\mathfrak{C}(\mathcal H(X)))$?
%
%Assume that for each combination of $(T,D,\tau)$, we have access to $N_{T,D,\tau}$ (denoted as $N$ in the following) independent identically distributed sequences $\bm x_1,\bm x_2, \dots, \bm x_N$ of white noise.
%Each of these sequences produces a point in the $H\times C$ plane, namely $(h_1,c_1), (h_2,c_2), \dots , (h_N,c_N)$.
%But let's remind how these points are produced.
%
%
%The Bandt-Pompe symbolization transforms each sequence of values $\bm x_i$ into a sequence of symbols $\bm \pi_i$.
%We build a histogram of proportions $\bm h_i$ with the sequence of symbols $\bm \pi_i$.
%Now we assume that $\bm h_i$ is an outcome of size $N$ of $X\sim\text{db}(\alpha,\beta)$, and we compute $\widehat\alpha_i,\widehat\beta_i$.
%In this way, the sequences $\bm x_1,\bm x_2, \dots, \bm x_N$ become histograms $\bm h_1, \bm h_2, \dots, \bm h_N$ and, on the one hand, they become points $(h_1,c_1), (h_2,c_2), \dots , (h_N,c_N)$ and estimates $(\widehat\alpha_1,\widehat\beta_1), (\widehat\alpha_2,\widehat\beta_2), \dots , (\widehat\alpha_N,\widehat\beta_N)$.
%We may then use $\text{db}(\overline{\widehat{\alpha}_i},\overline{\widehat{\beta}_i})$ distribution as a model for the histograms.
%From there, we may obtain an estimate of the distribution of $(\mathfrak{H},\mathfrak{C})$.
%
%The problem was knowing $\Pr\big((\mathfrak{H},\mathfrak{C})\in S\big)$, but $\mathfrak{H}=\mathfrak{H}(X)$  and $\mathfrak{C} = \mathfrak{C}(X)$, and we may assume that $X\sim\text{db}(\alpha,\beta)$, and we may estimate $(\alpha,\beta)$.
%The problem becomes finding the distribution of $\big(\mathfrak{H}(X),\mathfrak{C}(X)\big)$ under $X\sim\text{db}(\alpha,\beta)$.
